---
title: 'EDA: Long Form!'
author: "Data Science (STAT 301-1)"
subtitle: Anna Wagman
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    toc_float: yes
    highlight: tango
---

```{r setup, include=FALSE, message=FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Introduction:
As a subscriber to Hulu, AmazonPrime, and Disney+, I very excited to explore the similarities and differences between the three streaming services. In my EDA, I will explore the variables contained inside the individual data sets, and then compare specific variables across all three. Which service offers the most options? Does it differ depending on Movies vs TV Shows? Do Hulu, AmazonPrime, and Disney+ offer any of the same titles? How many? Which of these these services offer the most international streaming? In which countries? These are some of the questions I will be exploring and depicting throughout my Six Visualizations. The three data sets provide information on the titles available on Hulu, AmazonPrime, and Disney+ respectively. All three data sets include the same variables such as title, type, genre, rating, director, cast...etc

### Data Sources:
<https://www.kaggle.com/shivamb/hulu-movies-and-tv-shows/version/1.> <br>
<https://www.kaggle.com/shivamb/amazon-prime-movies-and-tv-shows> <br>
<https://www.kaggle.com/shivamb/disney-movies-and-tv-shows> <br>

### Data Cleaning: 
In my initial data cleaning, I renamed the column var = "listed_in" -> to "genre" for all three data sets and I removed the last column var = "description" that gave a lengthy description of the plot for each. Throughout my EDA, I use functions such as as.numeric, extract_numeric, round()...and many more in order to clean my data. For example, to work with var = "duration", I had to parse the numerical value from "duration" for each separate type = "Movie" and type = "TV Show". For type = "Movie", duration is given in minutes and so I used the "as.numeric()" function. For type = "TV Show", I had to extract the numerical value from duration given by "# Seasons" where '#' is an integer. I used the function "extract_numeric(duration)". I cleaned and recoded each variable in many different ways for each visualization, which is specifically outlined in my "data_cleaning.R" & "data_cleaning.Rmd" files which can be found in the main foler EDA_Wagman.


```{r, include = FALSE, message=FALSE, warning = FALSE}
#Load Packages
library(tidyverse)
library(janitor)
library(skimr)
library(ggplot2)
library(RColorBrewer)
library(lubridate)
library(cowplot)
library(viridis)
library(plyr)
library(purrr)
library(dbplyr)
library(ggmap)


#Load Data sets
hulu <- read_csv("data/processed/hulu_titles.csv")
disney <- read_csv("data/processed/disney_plus_titles.csv")
amazon <- read_csv("data/processed/amazon_prime_titles.csv")
```


# Visualization 1: Total TV Shows and Movies
### Q1. Which service has the most content? Is there a large disparity?

```{r, message = FALSE, include = FALSE, warning=FALSE}
#calculate the total of TV shows and Movies in each individual service
#make joint tibble of each service and total titles

#Hulu Movies + TV show
Hulu_MoviesvTV <- count(hulu, vars = "type")

#Hulu: total titles 
Hulu_Total <- sum(Hulu_MoviesvTV$freq)
Hulu_Total

#Disney+ Movies + TV show 
Disney_MoviesvTV <- count(disney, vars = "type")

#Disney+: total titles
Disney_Total <- sum(Disney_MoviesvTV$freq)
Disney_Total

#AmazonPrime Movies + TV shows
Amazon_MoviesvTV <- count(amazon, vars = "type")

#AmazonPrime: total titles
Amazon_Total <- sum(Amazon_MoviesvTV$freq)
Amazon_Total

#joint tibble with each service and the number of titles on each
TOTAL_COMPARISON <- tibble(
  service = c("Hulu", "Disney+", "Amazon Prime"), 
  total = c(Hulu_Total, Disney_Total, Amazon_Total)
  )
```

```{r, message = FALSE, include = FALSE, warning = FALSE}
##PLOT 1: Compare the total number of titles available on each streaming service

total_titles_plot <- ggplot(
  data = TOTAL_COMPARISON, #total titles per service
  aes(x = "", y = total, fill = service)) + #color by service
  geom_bar(stat = "identity") + 
  coord_polar("y", start = 0) + #pie chart
  geom_text(
    aes(label = total), #label amount of each
    position = position_stack(vjust = .5) #adjust labels
  ) +
  labs( #axes labels
    title = "Total titles on Hulu, Disney+, and AmazonPrime",
    x = NULL
  ) +
  scale_fill_manual(values = c("#FC6CB9", "#1EC2F9", "#5BDC70")) + #set color with specific HTML color codes
  theme_minimal()
```

```{r, message=FALSE, warning = FALSE}
#plot 1
total_titles_plot
```





# Visualization 2: Movies vs TV Shows on each service
### Q2. What is the distribution of Movies vs TV shows per service? 

```{r, include = FALSE, message = FALSE, warning = FALSE}
#Hulu Movies:
Hulu_Movies <- count(hulu, vars = "type")
Hulu_Movies_n <- as.numeric(Hulu_Movies[1, 2]) #extract first 2 columns
Hulu_Movies_n

#Disney+ Movies:
Disney_Movies <- count(disney, vars = "type")
Disney_Movies_n <- as.numeric(Disney_Movies[1, 2]) #extract first 2 columns
Disney_Movies_n

#AmazonPrime Movies
Amazon_Movies <- count(amazon, vars = "type")
Amazon_Movies_n <- as.numeric(Amazon_Movies[1, 2]) #extract first 2 columns
Amazon_Movies_n


##create tibble of total movies on each
Total_Movies <- tibble(
  service = c("Hulu", "Disney+", "Amazon Prime"), 
  total = c(Hulu_Movies_n, Disney_Movies_n, Amazon_Movies_n) 
  )
```


```{r, message=FALSE, include=FALSE, warning = FALSE}
##plot movies relative to each other
total_movies_plot <- ggplot(
  data = Total_Movies, #joint tibble movies on each
  aes(x = service, y = total)) + 
  geom_col(fill = c("#24C833", "#3A69C6", "#EA3BAA")) + #bar chart
  labs( #label axes
    title = "Number of Movies",
    y = "amount"
  ) +
  theme_minimal()
```

```{r, message=FALSE, include=FALSE, warning = FALSE}
#hulu: total TV Shows
hulu_TV <- hulu %>%
  filter(type == "TV Show") %>%
  group_by("title") %>%
  count()

#get number of rows = total
hulu_TV_total <- nrow(hulu_TV)

#disney: total TV Shows
disney_TV <- disney %>%
  filter(type == "TV Show") %>%
  group_by("title") %>%
  count() 

#number of rows
disney_TV_total <- nrow(disney_TV)

#amazon: total TV Shows
amazon_TV <- amazon %>%
  filter(type == "TV Show") %>%
  group_by("title") %>%
  count() 

#number of rows
amazon_TV_total <- nrow(amazon_TV)
```

```{r, message=FALSE, include = FALSE, warning = FALSE}
# create joint tibble of TV Shows on each
total_TVShows <- tibble(
  service = c("Hulu", "Disney+", "Amazon Prime"), 
  total = c(hulu_TV_total, disney_TV_total, amazon_TV_total)
  )

#plot TV Shows together
total_TVShows_plot <- ggplot(
  data = total_TVShows, 
  aes(x = service, y = total)) + 
  geom_col(fill = c("#5BDC70", "#1EC2F9", "#FC6CB9")) + #set colors
  labs(
    title = "Number of TV Shows",
    y = "amount"
  ) +
  theme_minimal()
```


```{r}
#plot side by side: Movies vs TV
plot_grid(total_movies_plot, total_TVShows_plot)
```


### 2b: Percentage of Movies vs TV Show on each

```{r, warning = FALSE, include = FALSE, message = FALSE}
#hulu percent break down
Hulu_Percentage_Movie <- (Hulu_Movies_n/Hulu_Total)*100 #divide by total and mult x 100
Hulu_Percentage_TV <- (hulu_TV_total/Hulu_Total)*100 #divide by total and mult x 100


Hulu_percent_tibble <- tibble(
  type = c("Movie", "TV Show"), 
  total = c(Hulu_Percentage_Movie, Hulu_Percentage_TV)
  )

Hulu_percent_plot <- ggplot(
  data = Hulu_percent_tibble, 
  aes(x = "", y = total, fill = type)) + #color by type: Movie vs TV
  geom_bar(stat = "identity") +
  coord_polar("y") + #pie chart
  labs(
    title = "Hulu: Movies vs TV",
    x = NULL
  ) +
  theme_minimal() +
  scale_fill_manual(values = c("#B567DC", "#F0E61F")) +
  geom_text(
    aes(label = round(total)), #label each (min) 
    position = position_stack(vjust = .5) #adjust labels
  )


#amazon percent break down
Amazon_Percentage_Movie <- (Amazon_Movies_n/Amazon_Total)*100
Amazon_Percentage_TV <- (amazon_TV_total/Amazon_Total)*100 


Amazon_percent_tibble <- tibble(
  type = c("Movie", "TV Show"), 
  total = c(Amazon_Percentage_Movie, Amazon_Percentage_TV)
  )

Amazon_percent_plot <- ggplot(
  data = Amazon_percent_tibble, 
  aes(x = "", y = total, fill = type)) + #color by type: Movie vs TV
  geom_bar(stat = "identity") +
  coord_polar("y") + #pie chart
  labs(
    title = "AmazonPrime: Movies vs TV",
    x = NULL
  ) +
  theme_minimal() +
  scale_fill_manual(values = c("#B567DC", "#F0E61F")) +
  geom_text(
    aes(label = round(total)), #label each (min) 
    position = position_stack(vjust = .5) #adjust labels
  )

#disney percent break down
Disney_Percentage_Movie <- (Disney_Movies_n/Disney_Total)*100
Disney_Percentage_TV <- (disney_TV_total/Disney_Total)*100 


Disney_percent_tibble <- tibble(
  type = c("Movie", "TV Show"), 
  total = c(Disney_Percentage_Movie, Disney_Percentage_TV)
  )

Disney_percent_plot <- ggplot(
  data = Disney_percent_tibble, 
  aes(x = "", y = total, fill = type)) + #color by type: Movie vs TV
  geom_bar(stat = "identity") +
  coord_polar("y") + #pie chart
  labs(
    title = "Disney+: Movies vs TV",
    x = NULL
  ) +
  theme_minimal() +
  scale_fill_manual(values = c("#B567DC", "#F0E61F")) +
  geom_text(
    aes(label = round(total)), #label each (min) 
    position = position_stack(vjust = .5) #adjust labels
  ) 
```


```{r, warning = FALSE, message = FALSE}
#plot percent Movie v TV of all 3 on one plot
plot_grid(Hulu_percent_plot, Amazon_percent_plot, Disney_percent_plot)
```



# Visualization 3: Movie Duration distribution
### Q3. Are movies on average the same length per service? 

```{r, message=FALSE, include=FALSE, warning = FALSE}
#convert duration to minutes for each service

#hulu movie minutes
hulu_movie_length <- hulu %>%
  select(type, duration, title) %>%
  filter(type == "Movie") %>% #filter by only movies
  drop_na() %>%
  #parse duration into an int:
  mutate(hulu_minutes = parse_number(duration)) #new var hulu_minutes


#amazon movie minutes
amazon_movie_length <- amazon %>%
  select(type, duration, title) %>%
  filter(type == "Movie") %>% #filter by only movies
  mutate(amazon_minutes = parse_number(duration)) #parse duration into an int

#disney movie minutes
disney_movie_length <- disney %>%
  select(type, duration, title) %>% 
  filter(type == "Movie") %>% #only movies
  drop_na() %>%
  mutate(disney_minutes = parse_number(duration)) #parse duration into an int
```


```{r, message = FALSE, include = FALSE, warning = FALSE}
#hulu movie duration distribution
hulu_duration_plot <- ggplot(
  data = hulu_movie_length, 
  mapping = aes(x = hulu_minutes)) +
  #histogram
  geom_histogram(binwidth = 10, #margin in minutes for each bar
                 color = "darkgreen", fill = "lightgreen") + 
  xlim(40, 200) + #set x-axes limits to scale plot
  theme_minimal() +
  labs(
    title = "Hulu: Average Movie length",
    x = "Duration in Minutes",
    y = "Amount of Movies"
  ) 
```


```{r, message = FALSE, include = FALSE, warning = FALSE}
#Plot 3b. AmazonPrime Movie Duration distribution

amazon_duration_plot <- ggplot(
  data = amazon_movie_length, 
  mapping = aes(x = amazon_minutes)) +
  #histogram
  geom_histogram(binwidth = 10, color = "red", fill = "pink") + 
  xlim(30, 200) + #set x-axes limits
  theme_minimal() +
  labs(
    title = "Average Movie length on AmazonPrime",
    x = "Duration in Minutes",
    y = "Amount of Movies"
  ) 
```


```{r, message = FALSE, warning = FALSE, include = FALSE}
#Plot 3c. Disney+ movie duration distribution

disney_duration_plot <- ggplot(
  data = disney_movie_length, 
  mapping = aes(x = disney_minutes)) +
  #histogram
  geom_histogram(binwidth = 10, #set margin for each bar in minutes
                 color = "darkblue", fill = "lightblue") + #histogram
  xlim(30, 200) + #set x-axes limits
  theme_minimal() +
  labs(
    title = "Average Movie length on Disney+",
    x = "Duration in Minutes",
    y = "Amount of Movies"
  ) 
```


```{r, message = FALSE, warning=FALSE}
#plot all three service_duration plots created above on one plot!
plot_grid(hulu_duration_plot, amazon_duration_plot, disney_duration_plot)
```




### 3b: Average (mean) of Movies Compared

```{r, message = FALSE, include=FALSE, warning = FALSE}
#calculate the mean duration each of a movie on each in minutes

#hulu average movie length (min)
mean_hulu_movies <- mean(hulu_movie_length$hulu_minutes)

#amazon average movie length (min)
mean_amazon_movies <- mean(amazon_movie_length$amazon_minutes)

#disney average movie length (min)
mean_disney_movies <- mean(disney_movie_length$disney_minutes)

#create joint tibble of the mean_service for all three
Movie_Averages <- tibble(
  service = c("Hulu", "Disney+", "Amazon Prime"), 
  total = c(mean_hulu_movies, mean_disney_movies, mean_amazon_movies)
  )
```

```{r, message = FALSE, include=FALSE, warning = FALSE}
#create bar plot of "mean movie duration" per service 
Movie_av_plot <- ggplot(
  data = Movie_Averages, 
  aes(x = service, y = total)) + 
  geom_col(fill = c("#5BDC70", "#1EC2F9", "#FC6CB9")) +
  labs(
    title = "Average Movie Length",
    y = "Average Length (min)"
  ) +
  theme_minimal() +
  geom_text(
    aes(label = round(total)), #label each (min) 
    position = position_stack(vjust = .5) #adjust labels
  ) 
```

```{r, message = FALSE, warning = FALSE}
Movie_av_plot
```



# Visualization 4: Titles Available on Multiple services

### Q4. Which services have overlap? How many titles do they have in common?

```{r, message = FALSE, include=FALSE, warning = FALSE}
#select all titles per service
hulu_titles <- hulu %>%
  select(title)

amazon_titles <- amazon %>%
  select(title)

disney_titles <- disney %>%
  select(title)

```

```{r, message = FALSE, warning = FALSE, include = FALSE}
# use inner_join between each combination of 2 of the services
# inner_join will join by "title" and only return titles in both

#hulu and amazon titles compared
hulu_and_amazon <- inner_join(hulu_titles, amazon_titles) 

#hulu and disney
disney_and_hulu <- inner_join(hulu_titles, disney_titles) 

#amazon and disney
amazon_and_disney <- inner_join(amazon_titles, disney_titles) 

#overlap between all three services
total_total_overlap <- inner_join(hulu_and_amazon, disney_titles)
```

```{r, warning=FALSE, message=FALSE, include = FALSE}
#rename columns and join all together:

names(hulu_and_amazon)[names(hulu_and_amazon) == 'title'] <- "hulu_amazon_titles"
names(disney_and_hulu)[names(disney_and_hulu) == 'title'] <- "disney_hulu_titles"
names(amazon_and_disney)[names(amazon_and_disney) == 'title'] <- "amazon_disney_titles"
```


```{r, warning=FALSE, message=FALSE, include=FALSE}
Hulu_Amazon_overlap <- nrow(hulu_and_amazon)
Disney_Hulu_overlap <- nrow(disney_and_hulu)
Amazon_Disney_overlap <- nrow(amazon_and_disney)


Total_Overlap <- tibble(
  services = c("Hulu and Amazon", "Disney and Hulu", "Amazon and Disney"),
  titles_overlap = c(Hulu_Amazon_overlap, Disney_Hulu_overlap, Amazon_Disney_overlap)
)
```
**Total titles availale on multiple services:**
```{r, warning=FALSE, message=FALSE}
Total_Overlap
```
**Titles availale on all 3 services:**

```{r, warning=FALSE, message=FALSE}
total_total_overlap
```

**As you can see, Hulu and AmazonPrime have the most overlap by a significant margin with 221 titles in common. The only title that all 3 services have in common is the movie "10 Things I Hate About You".**

# Visualization 5: Worldwide Streaming

### Q5. Which services offer international streaming? In which countries? How many titles are available in each? 

```{r, message = FALSE, include = FALSE, warning = FALSE}
# clean data
amazon_countries <- read_csv("data/processed/amazon_prime_titles.csv")  %>%
  select(country) %>% ## select the column var = "country"
  mutate(country = str_split(country,', ')) %>% ##read in multiple countries separated by commas  
  unnest(country) %>%
  group_by(country) %>% ##group by country name
  count() %>% ##count how many titles in each country
  arrange(desc(freq)) %>% ##order highest to lowest
  remove_empty() %>%
  head(20) ##return highest 20

#remove the first row which is country = NA...
amazon_countries_removeNA <- amazon_countries[-1, ]

#plot amazon countries (without first row = NA)
amazon_country_plot <- ggplot(amazon_countries_removeNA) +
  geom_col(fill = "pink", color = "red",
           aes(y = reorder(country, freq), x = freq)) +
  labs(
    title = "Amazon: Titles per Country",
    x = "Titles Available",
    y = "Country"
    ) +
  geom_label(aes(y = reorder(country,freq), x = freq, label = freq), size = 2) + #label freq per country
  theme_minimal()
```


```{r, message = FALSE, warning = FALSE, include=FALSE}
# Disney Countries
## split by "," for multiple countries
disney_countries <- read_csv("data/processed/disney_plus_titles.csv") %>%
  select(country) %>%
  mutate(country = str_split(country, ', ')) %>% ##read in multiple countries separated by commas 
  unnest(country) %>%
  group_by(country) %>% ##group by country
  count() %>% ##count titles per country
  arrange(desc(freq)) %>%
  head(20) ##select highest 20

#remove column country = NA
disney_removeNA <- disney_countries[-2, ]

#create plot of countries
disney_country_plot <- ggplot(disney_removeNA) +
  geom_col(fill = "lightblue", color = "darkblue",
           aes(y = reorder(country, freq), x = freq)) +
  labs(
    title = "Disney+: titles per Country",
    x = "Titles Available",
    y = "Country"
    ) +
  geom_label(aes(y = reorder(country,freq), x = freq, label = freq), size = 2) + #label freq per country
  theme_minimal()
```


```{r, message = FALSE, warning = FALSE}
plot_grid(amazon_country_plot, disney_country_plot)
```
<br>
**Hulu is only Available in the US and UK making for a very uninteresting graph, so I did not include Hulu. As shown above, the US has the most availability by a lot. The US has more titles than all of the other countries combined...**


# Visualization 6: Longest runnning TV Shows on each:
### Q6. What are the longest running TV Shows on each? How many seasons are there?

```{r, warning = FALSE, message=FALSE, include=FALSE}
# extract type = TV Show
# extract number from "2 Seasons"
# arrange descending, select top 10

##hulu tv seasons
hulu_seasons <- read_csv("data/processed/hulu_titles.csv") %>%
  filter(type == "TV Show") %>%
  select(title, duration) %>%
  mutate(duration = extract_numeric(duration)) %>% ##parse duration = "# Season" -> int; use extract_numeric() function
  arrange(desc(duration)) %>% ## arrange high -> low
  head(10) %>% ## return highest 10
  
  ggplot() +
  geom_col(fill = "lightgreen", color = "black",
    aes(y = reorder(title, duration), x = duration)) +
  labs(
    title = "Hulu: Most TV Seasons",
    x = "Amount of Seasons",
    y = "TV Show"
  ) +
  geom_label(aes(y = reorder(title, duration), x = duration, label = duration), size = 2.2) +
  theme_minimal()

##disney TV Show seasons
disney_seasons <- read_csv("data/processed/disney_plus_titles.csv") %>%
  filter(type == "TV Show") %>%
  select(title, duration) %>%
  mutate(duration = extract_numeric(duration)) %>% ##parse duration = "# Season" -> int
  arrange(desc(duration)) %>%
  head(10) %>%
  
  ggplot() +
  geom_col(fill = "lightblue", color = "black",
    aes(y = reorder(title, duration), x = duration)) +
  labs(
    title = "Longest Running TV Shows",
    x = "Disney+: Most TV Seasons",
    y = "TV Show"
  ) +
  geom_label(aes(y = reorder(title, duration), x = duration, label = duration), size = 2.2) +
  theme_minimal()


##amazon tv seasons
amazon_seasons <- read_csv("data/processed/amazon_prime_titles.csv") %>%
  filter(type == "TV Show") %>%
  select(title, duration) %>%
  mutate(duration = extract_numeric(duration)) %>% ##parse duration = "# Season" -> int
  arrange(desc(duration)) %>%
  head(10) 
  
amazon_seasons_plot <- ggplot(amazon_seasons) +
  geom_col(fill = "lightpink", color = "black",
           aes(y = reorder(title, duration), x = duration)) +
  labs(
    title = "Amazon: Most TV Seasons",
    x = "Amount of Seasons",
    y = "TV Show"
  ) +
  geom_label(aes(y = reorder(title, duration), x = duration, label = duration), size = 2.2) +
  theme_minimal()
```

```{r, warning = FALSE, message = FALSE}
hulu_seasons
disney_seasons
amazon_seasons_plot
```

<br>
**Hulu and AmazonPrime both have the TV Shows "Survivor" and "America's Top Model", both of which appear in their top 10 longest running shows. Fun fact: Survivor is my favorite TV Show ever and I have seen all 41 Seasons!**




